    ---
title: "Analyse the results of parameter scan"
output: html_notebook
---

Pay particular attention to the effect of window size and the standard deviation parameter. 


## Common Directories
```{r}
base_dir <- here::here()

data_dir <- file.path( base_dir, "Data" )

graph_dir <- file.path(data_dir, "Bedtools_Count")

#results_dir <- file.path(base_dir, "Results")
results_dir <- "/srv/scratch/treelab/SW_temporary_TermSeq_Results_of_run_pick_peak"

source_dir <- file.path(base_dir, "Source")

```


## Load libraries 
```{r}

if(!require(pacman)) {
  install.packages("pacman")
  libary(pacman)
}

p_load(tidyverse)
p_load(vroom)
p_load(here)
p_load(UpSetR)
p_load(plotly)
p_load(viridis)

source( file.path( source_dir, "Common", "common_functions.R") )
source( file.path( source_dir, "Common", "group_positions.R") )

sessionInfo()
```

## Common parameters
```{r}

use_log <- TRUE

```


## Directories and Files management
```{r}

project_output_dir <- file.path( results_dir, "Sites_B_subtilis" )

upset_counts_dir <- file.path(project_output_dir, "UpSet_Counts")
upset_ratios_dir <- file.path(project_output_dir, "UpSet_Ratios")
selected_params_dir <- file.path(project_output_dir, "Selected_Parameters")


if (use_log == TRUE ) {

  upset_counts_dir <- file.path(project_output_dir, "Log_UpSet_Counts")
  upset_ratios_dir <- file.path(project_output_dir, "Log_UpSet_Ratios")
  selected_params_dir <- file.path(project_output_dir, "Log_Selected_Parameters")

}

create_dir_if_not_exists(selected_params_dir)


## File with the optimal results 
selected_parameter_file <- file.path( selected_params_dir, 
                          "condition_vs_position_clusters_win8_sd7_cpm1_gw5.tab")

if( use_log == TRUE) {
  selected_parameter_file <- file.path( selected_params_dir, 
                          "condition_vs_position_clusters_win8_sd4_cpm0_gw5.tab")
  
}


```

```{r}
# Threshold number of replicates with reads for each genomic position for it to be included in further analysis
replicate_thresholds <- vroom::vroom( file.path( data_dir,
                         "Replicate_Thresholds",
                         "replicate_threshold_b_subtilis.tab" ))

# replicate_thresholds <- vroom::vroom( file.path( data_dir, 
#                          "Replicate_Thresholds", 
#                          "replicate_threshold_s_aureus_jkd6009.tab" ))

column_names_array <- replicate_thresholds %>% distinct(condition) %>% 
  arrange(condition) %>% pull(condition)  
```



```{r}
list_of_counts_files <- Sys.glob( file.path(upset_counts_dir, "upset_counts_win*_sd*.tab") )

```


```{r}

num_columns <- purrr::map(list_of_counts_files, vroom::vroom) %>% purrr::map_int(ncol)

counts_table <- purrr::map(list_of_counts_files[num_columns > 3], ~vroom::vroom(., id="filename") ) %>% 
                bind_rows() %>%
                mutate( filename = str_replace( filename, paste(upset_counts_dir, "/", sep=""), "" )) %>%
                mutate( window = str_replace( filename, ".*win(\\d+).*", "\\1"   )) %>%
                mutate( std_dev = str_replace( filename, ".*sd(\\d+).*", "\\1"   )) %>%
                dplyr::select( -filename )

counts_table


counts_table %>%
  distinct(window, std_dev)

# counts_table %>%
#   dplyr::filter(is.na(Amp) |
#                   is.na(Bac) |
#                   is.na(Cap) |
#                   is.na(Em) |
#                   is.na(Km) | 
#                   is.na(Lm) |
#                   is.na(Oflox))

```


## Find counts for each of the four conditions
```{r}
counts_table_clean <- counts_table %>%
  group_by( across( one_of( c(column_names_array, "cpm_cutoff", "distance_window", "window", "std_dev"))) ) %>%
  summarise( total_counts = sum(counts)) %>%
  ungroup() %>%
  mutate( window = as.integer(window),
          std_dev = as.integer(std_dev)) %>%
  arrange(  across( one_of( c(column_names_array, "window", "std_dev", "distance_window", "window"))) )

counts_table_clean
```

# Counts of peaks found in all four conditions - heat map
```{r}

dist5_cpm1_counts <- counts_table_clean %>%
  mutate( row_id = row_number()) %>%
  pivot_longer( cols=column_names_array,
                names_to="condition",
                values_to="is_identified") %>%
  mutate( is_identified = ifelse( is.na(is_identified), 0, is_identified)) %>%
  dplyr::filter( is_identified == 1) %>%
  arrange( cpm_cutoff, distance_window, window, std_dev, total_counts, row_id, condition) %>%
  group_by( row_id, cpm_cutoff, distance_window, window, std_dev, total_counts ) %>%
  dplyr::summarise( num_condition_identified = sum(is_identified)) %>%
  ungroup() %>%
  dplyr::filter( num_condition_identified >=4 ) %>%
  dplyr::filter( cpm_cutoff == 1 &
                   distance_window == 5)

min_window <- dist5_cpm1_counts %>% distinct(window) %>% pull(window) %>% min
max_window <- dist5_cpm1_counts %>% distinct(window) %>% pull(window) %>% max


dist5_cpm1_counts_heatmap <- dist5_cpm1_counts %>%
  mutate( window = factor(window, levels=seq(min_window, max_window,1) )) %>%
  ggplot( aes(window, std_dev, fill= total_counts)) + 
  geom_tile() +
  scale_fill_viridis() + 
  theme_bw()
    
dist5_cpm1_counts_heatmap

## don't know why ggplotly is not working
# ggplotly( dist5_cpm1_counts_heatmap)

```


# Total Number of peaks in all conditions - heat map
```{r}
total_counts_tbl <- counts_table_clean %>%
  group_by( distance_window, cpm_cutoff, window, std_dev) %>%
  summarise( total_counts=  sum( total_counts ) ) %>%
  ungroup()


dist5_cpm1_total_counts <- total_counts_tbl %>%
  dplyr::filter( cpm_cutoff == 1 &
                   distance_window == 5)


dist5_cpm1_total_counts_heatmap <- dist5_cpm1_total_counts %>%
  mutate( window = as.factor(window )) %>%
  ggplot( aes(window, std_dev, fill= total_counts)) + 
  geom_tile() +
  scale_fill_viridis() + 
  theme_bw()
  

ggplotly( dist5_cpm1_total_counts_heatmap)

```


# Ratio of Counts of peaks found in all four conditions / total number of peaks in all conditions, for all parameters

# Get a list of all of the files 
```{r}
list_of_ratios_files <- Sys.glob( file.path(upset_ratios_dir, "upset_ratios_win*_sd*.tab") )

```

# Read in the files,
```{r}
ratios_table_raw <- vroom::vroom(list_of_ratios_files, 
                                 id="filename")  # the full path of each file corresponding to each row
                                                 # stored in a column with the name 'filename'

```

## Clean up the table 
```{r}
ratios_table <- ratios_table_raw %>%
                ## replace the full path
                mutate( filename = str_replace( filename, paste(upset_counts_dir, "/", sep=""), "" )) %>%
                ## Extract the Window size used in peak picking
                mutate( window = str_replace( filename, ".*win(\\d+).*", "\\1"   )) %>%
                ## Extract the standard deviation used in peak picking
                mutate( std_dev = str_replace( filename, ".*sd(\\d+).*", "\\1"   )) %>%
                ## Remove the file name coumn 
                dplyr::select( -filename ) %>%
                ## Convert text column to integer column (as the window and std_dev were extracted from text)
                mutate( window = as.integer(window),
                        std_dev = as.integer(std_dev))  %>%
                ## Sort rows 
                arrange(  window, std_dev, distance_window, window)
```

## Ratios of groups of peaks found in all four conditions versus alll peaks in all conditions
This is the plot of the ratios of groups of peaks found in all four conditions versus total number of groups of peaks. You can see that window size of 8 with standard deviation of 7 is close to optimal.

2020-11-12 windows 5, standard deviation 3

```{r}
dist5_cpm1_ratios <- ratios_table  %>%
  dplyr::filter( cpm_cutoff == 1 &
                   distance_window == 5)



dist5_cpm1_ratios_heatmap <- dist5_cpm1_ratios  %>%
  mutate( window = as.factor(window )) %>%
  ggplot( aes(window, std_dev, fill= ratio)) + 
  geom_tile() +
  scale_fill_viridis() + 
  theme_bw()

ggplotly_ratios <- ggplotly(dist5_cpm1_ratios_heatmap)

ggplotly_ratios

htmlwidgets::saveWidget(ggplotly_ratios, "ggplotly_ratios.html")


```

## Read data for upset plot 
```{r}

selected_parameter_tbl <- vroom::vroom( selected_parameter_file )


```

```{r}

upset_freq <- upset( as.data.frame(selected_parameter_tbl), 
       order.by = "freq",
       decreasing = TRUE ) 

upset_freq

pdf( file.path(selected_params_dir, "upset_freq.pdf" ))
upset_freq
dev.off()


upset_degree <- upset( as.data.frame(selected_parameter_tbl), 
       order.by = "degree",
       decreasing = TRUE ) 

upset_degree

pdf( file.path(selected_params_dir, "upset_degree.pdf" ))
upset_degree
dev.off()


```

