---
title: "Extract term-site sequences for poly-U enrichment check "
output: html_notebook
---

## Common Directories
```{r}
base_dir <- here::here()

data_dir <- file.path( base_dir, "Data" )

graph_dir <- file.path(data_dir, "Bedtools_Count")

results_dir <- file.path(base_dir, "Results")

source_dir <- file.path(base_dir, "Source")

```

## Load libraries 
```{r}

if(!require(pacman)) {
  install.packages("pacman")
  libary(pacman)
}

p_load(tidyverse)
p_load(vroom)
p_load(here)
p_load(seqinr)
p_load(Biostrings)
p_load(ggseqlogo) # https://github.com/omarwagih/ggseqlogo # Draw sequence logo

source( file.path( source_dir, "Common", "common_functions.R") )
source( file.path( source_dir, "Common", "motif_analysis.R") )

```


## Common parameters
```{r}
  use_log <- TRUE
  window_length <- 8
  std_dev_cutoff <-  7
  replicate_cpm_cutoff <- 1
  distance_window <- 5
  
  
if(use_log==TRUE) {
  
  window_length <- 8
  std_dev_cutoff <-  4
  replicate_cpm_cutoff <- 1
  distance_window <- 5
  
  replicate_cpm_cutoff <- log2(replicate_cpm_cutoff)
}  

input_parameter_string <-  paste( "win", window_length,  "_",
                                  "sd", std_dev_cutoff,  
                                  sep="" ) 

full_parameter_string  <-  paste( "win", window_length,  "_",
                                  "sd", std_dev_cutoff,  "_",
                                  "cpm", replicate_cpm_cutoff, "_",
                                  "gw", distance_window,
                                  sep="" ) 

print(paste(  "window_length = ", window_length, ", ",
             "std_dev_cutoff = ", std_dev_cutoff,
             sep=""))

```

## Directories Management
```{r}

selected_params_dir <- file.path(results_dir, "Sites", "Selected_Parameters")

if (use_log == TRUE ) {
  selected_params_dir <- file.path(results_dir, "Sites", "Log_Selected_Parameters")
}

create_dir_if_not_exists(selected_params_dir)


```


```{r}

fasta_file <- file.path( data_dir, "Genomics/JDK6009/GCF_900607245.1_JKD6009_genomic.fna.gz" ) 
```


## Read the fasta file. 
```{r}

genome_seq <- read.fasta(file = fasta_file, 
  seqtype = c("DNA"))

genome_seq$NZ_LR027876.1

length(genome_seq$NZ_LR027876.1)

```



## Read master table 
```{r}

master_table <- vroom::vroom(  file.path( selected_params_dir, 
             paste0("master_table_", 
                    full_parameter_string, ".tab")))
```




# How to get a section of the genome, how to perform reverse complement then convert back to string
```{r}
genome_seq$NZ_LR027876.1[1000:1015]

temp_seq <- reverseComplement(DNAString(paste0(c(genome_seq$NZ_LR027876.1[1000:1015]), collapse="" )) )
temp_seq
toupper(as.character(temp_seq))
```



# Sequence Logos of the termination sites: 1. Termination sties in common among all four conditions 
Get 15 position including termination site 
```{r}
# Positive strand
positive_positions <- master_table %>%
  dplyr::filter( strand == "+") %>%
  dplyr::filter( cont == 1 &
                   linz == 1 &
                   tige == 1 &
                   vanco == 1) %>%
  pull(position)

positive_motif <- pos_strand_term_seq_substr(positive_positions, genome_seq$NZ_LR027876.1 )

ggplot() + 
  geom_logo(   positive_motif )  + 
  scale_x_continuous(breaks=c(1:16),
        labels=c(-15:0)) + 
  theme_logo()


# Negative strand
negative_positions <-  master_table %>%
  dplyr::filter( strand == "-") %>%
  dplyr::filter( cont == 1 &
                   linz == 1 &
                   tige == 1 &
                   vanco == 1) %>%
  pull(position)

negative_motif <- neg_strand_term_seq_substr(negative_positions, genome_seq$NZ_LR027876.1)

ggplot() + geom_logo(   negative_motif ) + theme_logo() + 
  scale_x_continuous(breaks=c(1:16),
        labels=c(-15:0)) + 
  theme_logo()
```      

## Positive and negative strand together
```{r}
 filtered_tbl <- master_table %>%
  dplyr::filter( cont == 1 &
                   linz == 1 &
                   tige == 1 &
                   vanco == 1)

term_seq_table <- term_seq_substr(filtered_tbl,  genome_seq$NZ_LR027876.1 )


intersect_all_conditions <- ggplot() + geom_logo( as.vector(term_seq_table$sequence )) + theme_logo() + 
  scale_x_continuous(breaks=c(1:16),
        labels=c(-15:0))

intersect_all_conditions

ggsave(filename=file.path(selected_params_dir, "seq_logs_intersect_all_conditions.pdf" ) , plot=intersect_all_conditions)

```



# Sequence Logos of the termination sites: 2. Termination sties from all conditions

## Positive and negative strand seperate
```{r}

# Positive strand
positive_positions <- master_table %>%
  dplyr::filter( strand == "+") %>%
  pull(position)


positive_motif <-  pos_strand_term_seq_substr(positive_positions, genome_seq$NZ_LR027876.1 )

ggplot() + geom_logo(   positive_motif ) + theme_logo() + 
  scale_x_continuous(breaks=c(1:16),
        labels=c(-15:0))

# Negative strand
negative_positions <-  master_table %>%
  dplyr::filter( strand == "-") %>%
  pull(position)


negative_motif <- neg_strand_term_seq_substr(negative_positions, genome_seq$NZ_LR027876.1)

ggplot() + geom_logo(   negative_motif ) + theme_logo() + 
  scale_x_continuous(breaks=c(1:16),
        labels=c(-15:0))
```

## Positive and negative strand together
```{r}
term_seq_table <- term_seq_substr(master_table,  genome_seq$NZ_LR027876.1 )


union_all_conditions <-  ggplot() + geom_logo(as.vector(term_seq_table$sequence ) ) + theme_logo() + 
  scale_x_continuous(breaks=c(1:16),
        labels=c(-15:0))

union_all_conditions

ggsave(filename=file.path(selected_params_dir, "seq_logs_union_all_conditions.pdf" ) , plot=union_all_conditions)
ggsave(filename=file.path(selected_params_dir, "seq_logs_union_all_conditions.png" ) , plot=union_all_conditions)


```
